---
title: "ESM244-Lab 4"
author: "Kira Archipov"
date: "2023-02-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)

library(tidyverse)
library(Metrics)
library(cowplot)
library(here)
library(kableExtra)
```

### Data 
```{r}
load(here("data", "fish_data.Rdata"))
```

### Single Model NLS

$$
S_t=B_{t+1}-B_t+C_t
$$

### Data Wrangling 
```{r}
surplus <- Fish_data %>% 
  group_by(stockid) %>% 
  select(stockid, year, TBbest, TCbest) %>% 
  drop_na() %>% 
  mutate(f_biomass = lead(TBbest)) %>% 
  mutate(surplus = f_biomass - TBbest + TCbest) %>% 
  drop_na()
```

### Plot 
```{r}
one_stock <- surplus %>% 
  filter(stockid == "COD1f-XIV")

ggplot(data = one_stock, aes(x = year, y = surplus)) +
  geom_point(size = 3, color = 'black') +
  theme_minimal()
```


### Step One: Select a Model 

Scour the literature, find what's out there and pick one

$$
\hat{S_t} = -e*MSY(\frac{B_t}{K}\ln(\frac{B_t}{K}))
$$
B = Biomass (g)
K = carrying capacity 
MSY = maximum sustainable yield 

## Creating Function 
```{r}
fox <- function(m, carry, biomass){
  out = -2.718*m*(biomass/carry)*log(biomass/carry)
  
  return(out)
}
```


### Step Two: Initial Guess 
```{r}
guess_vec = c(max(one_stock$TBbest)*0.37, 
              max(one_stock$TBbest))
```

### Step Three: Run NLS 
```{r}
one_stock_nls = nls(surplus ~ fox(m, carry, TBbest), 
                    data = one_stock, 
                    start = list(m = guess_vec[1], carry = guess_vec[2]),
                    trace = TRUE)

# will return our sum of sq errors, gradient, and then settles on new value
```

```{r}
one_stock_predict <- one_stock %>% 
  mutate(predict = predict(one_stock_nls, newdata = .))

ggplot(data = one_stock_predict) +
  geom_point(aes(x = year, y = surplus)) +
  geom_point(aes(x = year, y = predict), color = 'red')
```


```{r}
#adding controls, sometimes really helpful, sometimes (like this case) does not make a big difference, increases refinement

control_nls = nls(surplus ~ fox(m, carry, TBbest), 
                    data = one_stock, 
                    start = list(m = guess_vec[1], carry = guess_vec[2]),
                    trace = TRUE,
                  control = nls.control(tol = 2.1e-9)) # very precise tolerance 


```

### Using Purrr to run many NLS Models

We need to make our own function to put into purrr
```{r}
all_nls_fcn <- function(surplus_df){
  nls(surplus ~ fox(m, carry, TBbest), 
      data = surplus_df, 
      start = list(m = max(surplus_df$TBbest)*0.37, 
                   carry = max(surplus_df$TBbest)))
}

all_nls_fcn(one_stock)
```

## Run NLS for all species 
```{r}
fox_all <- surplus %>% 
  group_by(stockid) %>% 
  nest() %>% 
  mutate(nls_model = map(data, ~all_nls_fcn(.x))) %>% 
  mutate(predictions = map2(nls_model, data, ~predict(.x, newdata = .y))) %>% 
  mutate(rmse = map2_dbl(predictions, data, ~rmse(.x, .y$surplus)))

# want to compare how well the model fits internally 
```

### Graphs of twop 5 best fit model 
```{r}
plots <- fox_all %>% 
  arrange(rmse) %>% 
  head(5) %>% 
  mutate(graph = map2(data, predictions, 
                      ~ggplot()+
                        geom_point(data = .x, aes(x = .x$year, y = .x$surplus, 
                                                  color = 'Actual'))+
                        geom_point(aes(x = .x$year, y = .y, color = 'Predicted'))+
                        theme_minimal()+
                        xlab(' ')+
                        ylab('Surplus')+
                        scale_color_manual(name = 'Legend', breaks = c('Actual', 'Predicted'), 
                                           values = c('Actual' = 'black', 'Predicted' = 'Red'))))

#using cowplot to combine everything 

plot_list = plots$graph

cowplot::plot_grid(plotlist = plot_list, labels = c(plots$stockid, ""), hjust = -.05, vjust = 1)
```

```{r}
legend <- get_legend(plots$graph[[1]]) # pulls out legend 

for(i in 1:length(plots$graph)){
  plots$graph[[i]] <- plots$graph[[i]] + theme(legend.position = "none")
}

plot_legend <- plots$graph 

plot_legend[[6]] <- legend 

cowplot::plot_grid(plotlist = plot_legend, 
                   labels = c(plots$stockid, ""), hjust = -0.5, vjust = 1)

```













